---
title: "Global Situation of COVID19"
output: 
  flexdashboard::flex_dashboard:
    logo: img/logo.png
    favicon: img/favicon.png
    theme: flatly
    orientation: rows
    social: menu
    source_code: https://github.com/hossainlab/covid19-dashboard
    vertical_layout: fill
---

```{r setup, include=FALSE}
library(flexdashboard)

# Load packages 
library(tidyverse)
library(data.table)
library(DT)

# Color and maps 
library(plotly)
library(leaflet)
library(leafpop)
library(RColorBrewer)
library(glue)

# Dates
library(lubridate)
```


```{r, data-prep}
# Data 
library(coronavirus)
data("coronavirus")


# data prep for daily cases 
df_daily <- coronavirus |> 
  group_by(date, type) |> 
  summarise(total = sum(cases, na.rm = T), .groups = "drop") |> 
  pivot_wider(names_from = type, 
              values_from = total) |> 
  arrange(date) |> 
  ungroup() |> 
  dplyr::mutate(active = confirmed - death - recovery) |> 
  dplyr::mutate(confirmed_cum = cumsum(confirmed), 
         death_cum = cumsum(death),
         recovered_cum = cumsum(recovery),
         active_cum = cumsum(active))


# covid19 cases by country 
df_country <- coronavirus |> 
  group_by(country, type) |> 
  summarise(total = sum(cases, na.rm = T), .groups = "drop") |> 
  pivot_wider(names_from = type, 
              values_from = total) |> 
  dplyr::mutate(active = confirmed - death - recovery) |> 
  pivot_longer(cols = -country, 
               names_to = "type", 
               values_to = "total")


# for global statistics
df_world <- coronavirus |> 
  group_by(type) |> 
  summarise(total = sum(cases, na.rm = T), .groups = "drop") |>
  pivot_wider(names_from = type, 
              values_from = total) |> 
  mutate(active = confirmed - death - recovery)


# for leaflet map
df_map <- coronavirus  |> 
  filter(cases > 0)  |> 
  group_by(country, province, lat, long, type)  |> 
  summarise(cases = sum(cases), .groups = "drop")  |> 
  mutate(log_cases = 2 * log(cases))  |> 
  ungroup()

df_map.split <- df_map  |> 
  split(df_map$type)
```

```{r}
# custom color 
confirmed_color <- "#67000d"
active_color <-  "#1f77b4"
recovered_color <- "forestgreen"
death_color <- "red"
cfr_color <- "#660708"
discharged_pct_color <- "#e5383b"
```


# Overview 

Row
-----------------------------------------------------------------------
### confirmed {.value-box}
```{r}
confirmed <- format(df_world$confirmed, big.mark = ",")
valueBox(value = confirmed, 
         caption = "Total Confirmed Cases", 
         color = confirmed_color, 
         icon = "fa-virus")
```
### active {.value-box}

```{r}
active <- format(df_world$active, big.mark = ",")
valueBox(
    value = active,
    caption = "Active Cases",
    color = active_color, 
    icon = "fa-skull"
)
```


### recovered {.value-box}

```{r}
recovered <- format(df_world$recovery, big.mark = ",")

valueBox(
    value = recovered,
    caption = "Recovered Cases",
    color = recovered_color, 
    icon = "fa-check-circle"
)
```


### death {.value-box}

```{r}
death <- format(df_world$death, big.mark = ",")
valueBox(
    value = death, 
    caption = "Death Cases",
    color = death_color, 
    icon = "fa-exclamation-triangle"
)
```

### cfr {.value-box}

```{r}
cfr <- round((df_world$death / df_world$confirmed) * 100, 2)
valueBox(
    value = cfr,
    caption = "Case Fatality Rate (CFR)",
    color = cfr_color, 
    icon = "fa-heart-broken"
)
```

Row
-------------------------------
### Map 
```{r}

# Color pallete
pal <- colorFactor(c("grey", "red", "green"),
                   domain = c("confirmed", "death", "recovered"))



map_object <- leaflet()  |> 
  addProviderTiles(providers$CartoDB.DarkMatter) |> 
  setView(41.405559, 0.247590, zoom = 2)

names(df_map.split) |> 
  walk(function(coronavirus) {
    map_object <<- map_object |> 
      addCircleMarkers(data = df_map.split[[coronavirus]],
                       lng = ~long, lat = ~lat,
                       color = ~pal(type),
                       stroke = FALSE,
                       fillOpacity = 0.5,
                       radius = ~log_cases,
                       popup =  popupTable(df_map.split[[coronavirus]],
                                           feature.id = FALSE,
                                           row.numbers = FALSE,
                                           zcol=c("type","cases","country","province")),
                       group = coronavirus,
                       labelOptions = labelOptions(noHide = F,
                                                   direction = 'auto'))
  })

map_object |> 
  addLayersControl(
    overlayGroups = names(df_map.split),
    options = layersControlOptions(collapsed = FALSE)) |> 
  hideGroup(c("confirmed", "recovered"))
```


# Daily 
Column
-----------------------------------
### Daily Cumulative Cases
```{r}
# Daily Cumulative Cases
plot_ly(
  data = df_daily, 
  x = ~date, 
  y = ~active_cum, 
  name = "Active", 
  fillcolor = "blue", 
  type = "scatter", 
  mode = "none", 
  stackgroup = "one") |> 
  add_trace(y = ~confirmed_cum, 
            name = "Recovered", 
            fillcolor = "green") |> 
  add_trace(y = ~death_cum, 
            name = "Death", 
            fillcolor = "red") |> 
  layout(title = "", 
         xaxis = list(title = "Date", type = "date"), 
         yaxis = list(title = "Cumulative Number of Cases"), 
         legend = list(x = 0.1, y = 0.9), 
         hovermode = "compare")
```

### Confirmed Cases Distribution by Counrty
```{r}
# Plot the confirmed cases distribution by counrty with treemap plot
conf_df <- coronavirus |> 
  filter(type == "confirmed") |>
  group_by(country) |>
  summarise(total_cases = sum(cases)) |>
  arrange(-total_cases) |>
  mutate(parents = "Confirmed") |>
  ungroup() 

plot_ly(data = conf_df,
        type= "treemap",
        values = ~total_cases,
        labels= ~ country,
        parents=  ~parents,
        domain = list(column=0),
        name = "Confirmed",
        textinfo="label+value+percent parent")
```

# Top20
Column{.tabset}
-----------------------------------
```{r}
# top 20 
top_20 <- df_country |> 
  group_by(country, type) |> 
  pivot_wider(names_from = type, values_from = total) |> 
  arrange(-confirmed) |> 
  ungroup() |> 
  top_n(20)


# define 20 colors with colorRampPalette
ncountry <- 20 
country_color <- colorRampPalette(brewer.pal(10, "RdYlGn"))(ncountry)
```


### Top 20 Countries for Confirmed Caases 
```{r}
# confirmed cases 
plot_ly(data = top_20, 
        x = ~country, 
        y = ~confirmed, 
        type = "bar",
        marker = list(color = country_color)) |> 
  layout(title = "", 
         xaxis = list(title = "Country", categoryorder = "total descending"), 
         yaxis = list(title = "Confirmed Cases"))

```

### Top 20 Countries for Active Caases 
```{r}
# active cases 
plot_ly(data = top_20, 
        x = ~country, 
        y = ~active, 
        type = "bar", 
        marker = list(color = country_color)) |> 
  layout(title = "", 
         xaxis = list(title = "Country", categoryorder = "total descending"), 
         yaxis = list(title = "Active Cases"))

```


### Top 20 Countries for Recovered Caases 
```{r}
# recovered cases 
plot_ly(data = top_20, 
        x = ~country, 
        y = ~recovery, 
        type = "bar", 
        marker = list(color = country_color)) |> 
  layout(title = "", 
         xaxis = list(title = "Country", categoryorder = "total descending"), 
         yaxis = list(title = "Recovered Cases"))
```

### Top 20 Countries for Death Caases 
```{r}

# death cases 
plot_ly(data = top_20, 
        x = ~country, 
        y = ~death, 
        type = "bar", 
        marker = list(color = country_color)) |> 
  layout(title = "", 
         xaxis = list(title = "Country", categoryorder = "total descending"), 
         yaxis = list(title = "Death Cases"))
```

# Vaccination  
```{r}

# Vaccination Data
df_vax <- read.csv("data/vaccinations.csv")


# daily vaccinations
df_vax_daily <- df_vax |> 
  select(date, location, daily_vaccinations) |> 
  group_by(location, date) |> 
  summarise(total = sum(daily_vaccinations), .groups="drop") |> 
  filter(location != "World") |> 
  arrange(-total)|> 
  pivot_wider(names_from = date, values_from = total) |> 
  slice(1:20) |> 
  pivot_longer(cols = -location, 
               names_to = "date", 
               values_to = "total") |> 
  mutate(date = ymd(date)) 



# world statistics
df_world_vax <- df_vax |> 
  select(location,
         total_vaccinations,
         people_vaccinated,
         people_fully_vaccinated,
  ) 
# world statistics
df_world_vax <- df_vax |> 
  select(location,
         total_vaccinations,
         people_vaccinated,
         people_fully_vaccinated,
  ) |> 
  filter(location == "World") |> 
  summarise(total_vax = max(total_vaccinations, na.rm = T),
            ppl_once = max(people_vaccinated, na.rm = T),
            fully_vax = max(people_fully_vaccinated, na.rm = T)) |> 
  mutate(across(where(is.numeric), scales::comma))


# total vaccines administered
df_vax_total <- df_vax |> 
  select(location, iso_code, total_vaccinations) |> 
  group_by(location) |> 
  filter(total_vaccinations == max(total_vaccinations, na.rm = T), location != "World") 
```


Row
-----------------------------------------------------------------------

### Total Doses {.value-box}

```{r}
valueBox(value = df_world_vax$total_vax,
         caption = "Total Doses Administered",
         color = "#99d8c9")
```

### At least one {.value-box}

```{r}
valueBox(value = df_world_vax$ppl_once,
         caption = "No. of people vaccinated (at least one dose)",
         color = "#66c2a4")
```

### People Fully vaccinated {.value-box}

```{r}
valueBox(value = df_world_vax$fully_vax, 
         caption = "No. of people fully vaccinated", 
         color = "#41ae76")
```

Row{.tabset}
-----------------------------------------------------------------------

### Map 
```{r}

# figure out how to use iso code to create leaflet maps
# https://www.r-graph-gallery.com/183-choropleth-map-with-leaflet.html

# Plotly choroplet map
# https://plotly.com/r/choropleth-maps/
# light grey boundaries
l <- list(color = toRGB("grey"), width = 0.5)

# specify map projection/options
g <- list(
  showframe = FALSE,
  showcoastlines = FALSE,
  projection = list(type = 'Mercator')
)

# add trace
vax_map <- plot_geo(df_vax_total)
vax_map <- vax_map  |> 
  add_trace(
  z = ~total_vaccinations,
  color = ~total_vaccinations,
  colors = 'Greens',
  text = ~location,
  locations = ~iso_code,
  marker = list(line = l)
)

# add title to colorbar
vax_map <- vax_map  |> 
  colorbar(title = 'Total vaccinations (M)')

# add title and subtitle
vax_map <- vax_map  |> 
  layout(title = 'Total COVID-19 Vaccination Doses Administered<br>Source:<a href="https://github.com/owid/covid-19-data/tree/master/public/data/vaccinations">Our World in Data</a>', geo = g)

vax_map
```

### Top 20 Countries 
```{r}

# Top 20
# define 20 colors with colorRampPalette
ncountry_vaccination <- 20 
country_color_vaccination <- colorRampPalette(brewer.pal(10, "RdYlGn"))(ncountry_vaccination)


# for limits of graph
max = arrange(df_vax_total, -total_vaccinations)$total_vaccinations[1]

vax_total <- df_vax_total  |> 
  arrange(-total_vaccinations) |> 
  ungroup() |> 
  slice(1:20) |> 
  mutate(location = fct_reorder(location, total_vaccinations))


plot_ly(data = vax_total, 
        x = ~location, 
        y = ~total_vaccinations, 
        type = "bar", 
        marker = list(color = country_color_vaccination)) |> 
  layout(title = "", 
         xaxis = list(title = "Country", categoryorder = "total descending"), 
         yaxis = list(title = "Total Vaccinations"))

```



### Data 
```{r}
df_vax_total |> 
  datatable(options = list(
    bPaginate = TRUE,
    scrollX = FALSE,
    columnDefs = list(list(className = 'dt-center', targets = "_all"))
  ),
  rownames=FALSE,
  caption = 'Vaccination Data',
  style = 'bootstrap', 
  class = 'cell-border stripe')
```


# About 
This dashboard provides an overview of the COVID-19 (2019-nCoV) epidemic. This dashboard was built with R using Rmarkdown and the Flexdashboard package. The Github Repository for this dashboard available [here](https://github.com/hossainlab/covid19-dashboard)

**Data**

The source of data for this dashboard is from the R package [coronavirus](https://github.com/RamiKrispin/coronavirus). With the raw data pulled from the [COVID-19 Data Repository by the Center for Systems Science and Engineering (CSSE) at Johns Hopkins University](https://github.com/CSSEGISandData/COVID-19)

The source of vaccination data can be found at the Our World in Data [vaccination github repo](https://github.com/owid/covid-19-data/tree/master/public/data/vaccinations)

**Packages**

* Dashboard interface - `flexdashboard`
* Visualization - `plotly`, `RColorBrewer`, `ggplot2`
* Data Import - `data.table`
* Data manipulation - `dplyr`, `tidyr`, `purrr`, `forcats`
* Mapping - `leaflet`, `leafpop` 
* Strings - `glue`
* Data Source - `coronavirus`
* Data display - `DT`
* Dates - `lubridate`
* Scales - `scales`
